## setup

```{r}
#| eval: true
#| label: capeml

devtools::load_all("~/localRepos/capeml/")
```

```{r}
#| eval: true
#| label: connection

source("~/Documents/localSettings/pg_local.R")
pg <- pg_local_connect("caplter")
```

```{r}
#| eval: true
#| label: queries

source("esca_sql_queries.R")
```


## human indicators

```{r}
#| eval: true
#| label: human_indicators

human_indicators <- get_human_indicators('survey200')

human_indicators <- human_indicators |>
  dplyr::mutate(
    human_presence_of_path    = as.factor(human_presence_of_path),
    human_footprints          = as.factor(human_footprints),
    human_bike_tracks         = as.factor(human_bike_tracks),
    human_off_road_vehicle    = as.factor(human_off_road_vehicle),
    human_small_litter        = as.factor(human_small_litter),
    human_dumped_trash_bags   = as.factor(human_dumped_trash_bags),
    human_abandoned_vehicles  = as.factor(human_abandoned_vehicles),
    human_graffiti            = as.factor(human_graffiti),
    human_injured_plants      = as.factor(human_injured_plants),
    human_informal_play       = as.factor(human_informal_play),
    human_informal_recreation = as.factor(human_informal_recreation),
    human_informal_living     = as.factor(human_informal_living),
    human_sports_equipment    = as.factor(human_sports_equipment),
    human_social_class        = as.factor(human_social_class)
  )

try({
  capeml::write_attributes(human_indicators, overwrite = FALSE)
  capeml::write_factors(human_indicators, overwrite = FALSE)
})
```


## landscape irrigation

```{r}
#| eval: true
#| label: landscape_irrigation

landscape_irrigation <- get_landscape_irrigation("survey200")

landscape_irrigation <- landscape_irrigation |>
  dplyr::mutate(
    appears_maintained              = as.factor(appears_maintained),
    appears_professional            = as.factor(appears_professional),
    appears_healthy                 = as.factor(appears_healthy),
    appears_injured                 = as.factor(appears_injured),
    presence_of_open_ground         = as.factor(presence_of_open_ground),
    presence_of_trees               = as.factor(presence_of_trees),
    presence_of_shrubs              = as.factor(presence_of_shrubs),
    presence_of_cacti_succ          = as.factor(presence_of_cacti_succ),
    presence_of_lawn                = as.factor(presence_of_lawn),
    presence_of_herbaceous_ground   = as.factor(presence_of_herbaceous_ground),
    presence_of_other               = as.factor(presence_of_other),
    presence_of_hand_water          = as.factor(presence_of_hand_water),
    presence_of_drip_water          = as.factor(presence_of_drip_water),
    presence_of_overhead_water      = as.factor(presence_of_overhead_water),
    presence_of_flood_water         = as.factor(presence_of_flood_water),
    presence_of_subterranean_water  = as.factor(presence_of_subterranean_water),
    presence_of_no_water            = as.factor(presence_of_no_water),
    presence_of_pervious_irrigation = as.factor(presence_of_pervious_irrigation)
  )

try({
  capeml::write_attributes(landscape_irrigation, overwrite = FALSE)
  capeml::write_factors(landscape_irrigation, overwrite = FALSE)
})
```


## annuals

```{r}
#| eval: true
#| label: annuals

annuals <- get_annuals() |>
  dplyr::mutate(
    vegetation_rope_length = dplyr::case_when(
      vegetation_rope_length > 60 ~ 60,
      TRUE ~ vegetation_rope_length
    )
  )

try({
  capeml::write_attributes(annuals, overwrite = FALSE)
})
```


## shrub surveys

Some 2015 counts appear to be duplicated as evidenced by identical values in
each quadrant for each taxa. One instance from each of the duplicates,
documented below, is excluded from the query. Note that this is not a database
or query issue as the duplication is reflected in the tablet output. Because the
issue cannot be resolved unequivocally, the records are removed from the query
rather than removed from the database.

```
site_code ,samp_date  ,vegetation_scientific_name                             ,vegetation_sample_id
AB10      ,2015-04-09 ,Acacia constricta                                      ,               28952
AB10      ,2015-04-09 ,Acacia greggii                                         ,               28951
AB10      ,2015-04-09 ,Calliandra eriophylla                                  ,               28949
AB10      ,2015-04-09 ,Celtis pallida                                         ,               28950
AB10      ,2015-04-09 ,Cylindropuntia acanthocarpa                            ,               28944
AB10      ,2015-04-09 ,Krameria grayi                                         ,               28946
AB10      ,2015-04-09 ,Leucophyllum laevigatum                                ,               28943
AB10      ,2015-04-09 ,Phoradendron californicum                              ,               28953
AB10      ,2015-04-09 ,Senna covesii                                          ,               28947
AB10      ,2015-04-09 ,Sphaeralcea ambigua                                    ,               28948
AB10      ,2015-04-09 ,Viguiera deltoidea var. parishii (Greene) Vasey & Rose ,               28945
F8        ,2015-04-06 ,Ambrosia deltoidea                                     ,               28587
F8        ,2015-04-06 ,Larrea tridentata                                      ,               28586
V7        ,2015-04-08 ,Acacia greggii                                         ,               28028
V7        ,2015-04-08 ,Ambrosia deltoidea                                     ,               28025
V7        ,2015-04-08 ,Baccharis                                              ,               28031
V7        ,2015-04-08 ,Cylindropuntia acanthocarpa                            ,               28027
V7        ,2015-04-08 ,Ferocactus cylindraceus                                ,               28023
V7        ,2015-04-08 ,Larrea tridentata                                      ,               28026
V7        ,2015-04-08 ,Lycium                                                 ,               28024
V7        ,2015-04-08 ,Lycium                                                 ,               29344
V7        ,2015-04-08 ,Lycium andersonii                                      ,               28032
V7        ,2015-04-08 ,Ziziphus obtusifolia                                   ,               28029
```

```{r}
#| eval: true
#| label: shrub_surveys

shrubs_surveys <- get_shrub_surveys() |>
  dplyr::mutate(
    vegetation_shape_code          = as.factor(vegetation_shape_code),
    vegetation_classification_code = as.factor(vegetation_classification_code)
  ) |>
  dplyr::select(
    sample_date,
    site_code,
    vegetation_scientific_name,
    vegetation_classification_code,
    vegetation_shape_code,
    height,
    width_ns,
    width_ew
  )

try({
  capeml::write_attributes(shrubs_surveys, overwrite = FALSE)
  capeml::write_factors(shrubs_surveys, overwrite = FALSE)
})

# dplyr::mutate(
#   calc = dplyr::case_when(
#     height_degree_down < 0 ~ (tan(height_degree_up) - tan(height_degree_down)) * height_distance,
#     height_degree_down > 0 ~ (tan(height_degree_up) - tan(height_degree_down)) * height_distance
#   )

```


## trees

```{r}
#| eval: true
#| label: trees

trees <- get_trees("survey200") |>
  dplyr::mutate(
    vegetation_shape_code          = as.factor(vegetation_shape_code),
    vegetation_classification_code = as.factor(vegetation_classification_code),
    canopy_condition               = as.factor(canopy_condition)
  ) |>
  dplyr::select(
    sample_date,
    site_code,
    vegetation_scientific_name,
    vegetation_classification_code,
    vegetation_shape_code,
    height = height_in_m,
    crown_width_ns,
    crown_width_ew,
    stem_diameter,
    stem_height = stem_diameter_at,
    stem_count,
    missing_branches,
    canopy_condition
  )

try({
  capeml::write_attributes(trees, overwrite = FALSE)
  capeml::write_factors(trees, overwrite = FALSE)
})

trees <- get_trees("survey200") |>
  dplyr::mutate(
    vegetation_shape_code          = as.factor(vegetation_shape_code),
    vegetation_classification_code = as.factor(vegetation_classification_code),
    canopy_condition               = as.factor(canopy_condition)
  ) |>
dplyr::mutate(
  calc = dplyr::case_when(
    !is.na(height_degree_down) & !is.na(height_degree_down) & !is.na(height_degree_down) ~ (tan(height_distance) - tan(height_degree_down)) * height_distance,
    # height_degree_down < 0 ~ (tan(height_degree_up) - tan(height_degree_down)) * height_distance,
    # height_degree_down > 0 ~ (tan(height_degree_up) - tan(height_degree_down)) * height_distance
  )
)
```

## perennials

```{r}
#| eval: true
#| label: perennials

number_perennials <- get_perennials()

try({
  capeml::write_attributes(number_perennials, overwrite = FALSE)
})
```